---
title: "Recipes for some analyses modifications"
subtitle: "Where we present some piece of codes to modify the bioinformatic pipeline"
author: Adrien Taudière
date: last-modified
execute: 
  eval: false
format: 
  html:     
    code-fold: false
# bibliography: bibliography.bib
# link-citations: true
---

See also [Easy16S](https://github.com/YongxinLiu/EasyAmplicon/blob/master/pipeline_en.sh) for inspiration in bash. 

## Change the parameter to merged forward and reverse sequences 

1. Play with option min Overlap (default to 12) and maxMismach (default to 0). For example, replace the `merged_seq` 
targets with the following code to allow 1 mismatch and a minimum overlap of 8.

```{r, eval=FALSE}
tar_target(
    merged_seq,
    mergePairs(
      dadaF = ddF,
      dadaR = ddR,
      derepF = derep_fs,
      derepR = derep_rs,
      minOverlap = 8,
      maxMismatch = 1)
    ),
    format = "qs"
  )
)

```



## Forward only pipeline


1. Replace all the "paired end" area with the following code

```{r, eval=FALSE}
  ##> Remove primers
  tar_target(
    cutadapt,
    cutadapt_remove_primers(
      path_to_fastq = here("data/data_raw/rawseq/"),
      primer_fw = fw_primer_sequences,
      folder_output = here("data/data_intermediate/seq_wo_primers/"),
      args_before_cutadapt = "source ~/miniforge3/etc/profile.d/conda.sh && conda activate cutadaptenv && "
    )
  ),
  tar_target(data_raw, {
    cutadapt
    list_fastq_files(path = here::here("data/data_intermediate/seq_wo_primers/"),
                     paired_end = FALSE)
  }),

  ##> Classical dada2 pipeline
  tar_target(data_fnfs, data_raw$fnfs),
  ### Pre-filtered data with low stringency
  tar_target(
    filtered,
    filter_trim(
      output_fw = paste(
        getwd(),
        here("/data/data_intermediate/filterAndTrim_fwd"),
        sep = ""
      ),
      rev = data_fnrs,
      multithread = n_threads,
      compress = TRUE
    )
  ),

  ### Dereplicate fastq files
  tar_target(derep_fs, derepFastq(filtered[[1]]), format = "qs"),
  tar_target(derep_rs, derepFastq(filtered[[2]]), format = "qs"),
  ### Learns the error rates
  tar_target(err_fs, learnErrors(derep_fs, multithread = 4), format = "qs"),
  tar_target(err_rs, learnErrors(derep_rs, multithread = 4), format = "qs"),
  ### Make amplicon sequence variants
  tar_target(ddF, dada(derep_fs, err_fs, multithread = 4), format = "qs"),
  tar_target(ddR, dada(derep_rs, err_rs, multithread = 4), format = "qs"),
  ### Build a a table of ASV x Samples
  tar_target(seq_tab, makeSequenceTable(ddF)),

```



## Add a second taxonomic assignation using a different database or algorithm

1. Add a new database file (*fasta*) in `data/data_raw/refseq`
1. Copy and complete with good names the two targets below
1. Rename the targets by using the new name (e.g. `data_phyloseq_newDB`) instead of `data_phyloseq` in the subsequent targets


```{r, eval=FALSE}
[...]
tar_target(
    name = file_refseq_taxo2,
    command = "data/data_raw/refseq/XXX",
    format = "file"
)

[...]

tar_target(
  data_phyloseq_newDB,
  add_new_taxonomy_pq(
    data_phyloseq,
    file_refseq_taxo2,
    suffix = "PR2",
    taxLevels = c(
      "Kingdom",
      "Supergroup",
      "Division",
      "Subdivision",
      "Class",
      "Order",
      "Family",
      "Genus",
      "Species"
    )
  )
)
```


## Filter by % sequence (whith blast)

```{r, eval=FALSE}
tar_target(d_blast,
    filter_taxa_blast(
      data_phyloseq,
      fasta_for_db = paste0(here::here(), "/", file_refseq_taxo),
      nproc = 4
    )
)
```

## Add funguild informations for Fungi

```{r, eval=FALSE}
tar_target(d_funguild, 
  MiscMetabar::add_funguild_info(data_phyloseq)
)
```



## Add Protax informations for Bacteria


::: {.content-hidden}

## FAPROTAX analysis

### Assign functionnality 

```{r}
# Inspiration from https://forum.qiime2.org/t/exporting-otu-table-from-phyloseq-into-either-biom-or-text-format/19103/7
write_biom_csv <- function(physeq, file, sep = "; ") {
  ps <- phyloseq::taxa_sums(physeq) |> 
    as.data.frame()
 
  ps |>  tibble::rownames_to_column("#OTU ID") |> 
    left_join(phyloseq::tax_table(physeq) |>  
                as.data.frame() |> 
                tibble::rownames_to_column("#OTU ID") |>  
                tidyr::unite("taxonomy", !`#OTU ID`, sep = sep)) -> phyloseq_biom
  
  write.table(phyloseq_biom, file = file, sep = "\t", quote = FALSE, row.names = FALSE)
}

```


```{r}
here::i_am("report_16S.qmd")
tax_tab_path <- here::here("output/faprotax_tax_table.csv")
FAPROTAX_path <- here::here("code/FAPROTAX_1.2.7")
d_vs@tax_table <- tax_table(cbind(d_vs@tax_table,   "Genus_species" = paste(d_vs@tax_table[,"Genus"], 
                                                      d_vs@tax_table[,"Species"]) ))
write.table(subset_taxa(d_vs, !is.na(Species))@tax_table, tax_tab_path, sep = ";")
write_biom_csv(d_vs, tax_tab_path)
dir.create("output/FAPROTAX")

# system(pip install biom-format)
system(paste0("python3 ", FAPROTAX_path, "/collapse_table.py", " -i ", tax_tab_path, " -o output/FAPROTAX/func_table.csv -g ", FAPROTAX_path, '/FAPROTAX.txt -r output/FAPROTAX/report.txt -s output/FAPROTAX/output_subtables/ -f -c "#" -d "taxonomy" --omit_columns 0 --column_names_are_in last_comment_line -n columns_after_collapsing -v --out_groups2records_table output/FAPROTAX/faprotax_result_OTU.tsv'))

func_tab <- read.table("output/FAPROTAX/faprotax_result_OTU.tsv", sep="\t", header = TRUE)
colnames(func_tab) <- gsub("FPTax_record", "record_faprotax", paste0("FPTax_", colnames(func_tab)))


tax_tab <-
    as.matrix(cbind(
      tax_table(d_vs),
      func_tab
    ))
d_vs@tax_table <- tax_table(tax_tab)
```

### Functionnal analysis

```{r}
func_group <- colSums(apply(d_vs@tax_table[,grepl("FPTax",colnames(d_vs@tax_table))], 2, as.numeric))
func_group <- func_group[func_group>0]

DT::datatable(as.data.frame(sort(func_group, decreasing = T)))

func_group_more_than20 <- func_group[func_group>20]

d_vs_Treatment <- speedyseq::merge_samples2(d_vs, "Treatment")
```


```{r}
psm <- psmelt(d_vs) |> 
  filter(Abundance > 0) |> 
  mutate(across(all_of(starts_with("FPTax_")), as.numeric))

psm |> 
  group_by(Sample) |>  
  summarise("sumFPTax_chemoheterotrophy" = sum(FPTax_chemoheterotrophy))
   
ggplot(psm, aes(x = Treatment, y = FPTax_chemoheterotrophy, fill = Sample)) +
  geom_bar(stat = "identity")

ggplot(psm, aes(x = Treatment, y = FPTax_chemoheterotrophy)) + geom_boxplot()
  
psm2 <- psm |>  
  group_by(Sample) |> 
  summarise(across(all_of(starts_with("FPTax_")), sum))  |>  
  tidyr::pivot_longer(all_of(starts_with("FPTax_")),
                      names_to = "Function",
                      values_to = "Nb_ASV")

psm2_na <- psm |>
  group_by(Sample) |>
  summarise(across(all_of(starts_with("FPTax_")), function(x) {
    sum(x == 0)
  }))  |>
  tidyr::pivot_longer(all_of(starts_with("FPTax_")),
                      names_to = "Function",
                      values_to = "Nb_ASV_NA")

psm2_nbseq <- psm |>
  group_by(Sample) |>
  summarise(across(all_of(starts_with("FPTax_")), function(x) {
    sum(x * Abundance)
  }))  |>
  tidyr::pivot_longer(all_of(starts_with("FPTax_")),
                      names_to = "Function",
                      values_to = "Nb_seq")

psm3 <-  psm2 |> 
  filter(Function %in% names(func_group_more_than20)) 
psm3_na <-  psm2_na |> 
  filter(Function %in% names(func_group_more_than20)) 
psm3_nbseq <-  psm2_nbseq |> 
  filter(Function %in% names(func_group_more_than20)) 


psm_4 <- cbind(psm3, 
                    "Nb_ASV_NA"=psm3_na$Nb_ASV_NA, 
                    "Nb_seq"=psm3_nbseq$Nb_seq)

psm_4 <- psm_4 |> 
  mutate("Prop_ASV" = Nb_ASV/(Nb_ASV + Nb_ASV_NA))

psm_4$Treatment <- d_vs@sam_data$Treatment[match(psm_4$Sample, sample_names(d_vs))]

ggplot(psm_4, aes(y = Function, x = Nb_ASV, fill= Treatment)) +
  geom_boxplot()

ggplot(psm_4, aes(y = Function, x = Nb_seq, fill= Treatment)) +
  geom_boxplot()

ggstatsplot::ggbetweenstats(psm_4 |> filter(Function == "FPTax_chemoheterotrophy"), 
                            Treatment, Nb_ASV)
```



```{r}
psm_treat <- psmelt(d_vs_Treatment) |> 
  filter(Abundance > 0) |> 
  mutate(across(all_of(starts_with("FPTax_")), as.numeric))

psm_treat2 <- psm_treat |>  
  group_by(Treatment) |> 
  summarise(across(all_of(starts_with("FPTax_")), sum)) |> 
  group_by(Treatment)   |>  
  tidyr::pivot_longer(all_of(starts_with("FPTax_")),
                      names_to = "Function",
                      values_to = "Nb_ASV")

psm_treat2_na <- psm_treat |>  
  group_by(Treatment) |> 
  summarise(across(all_of(starts_with("FPTax_")), function(x){
    sum(x == 0) 
  })) |> 
  group_by(Treatment)   |>  
  tidyr::pivot_longer(all_of(starts_with("FPTax_")),
                      names_to = "Function",
                      values_to = "Nb_ASV_NA")

psm_treat2_seq <- psm_treat |>  
  group_by(Treatment) |> 
  summarise(across(all_of(starts_with("FPTax_")), function(x){
   sum(x*Abundance)
  })) |> 
  group_by(Treatment)   |>  
  tidyr::pivot_longer(all_of(starts_with("FPTax_")),
                      names_to = "Function",
                      values_to = "Nb_seq")


psm_treat3 <-  psm_treat2 |> 
  filter(Function %in% names(func_group_more_than20))
psm_treat3_na <-  psm_treat2_na |> 
  filter(Function %in% names(func_group_more_than20))
psm_treat3_seq <-  psm_treat2_seq |> 
  filter(Function %in% names(func_group_more_than20))

psm_treat4 <- cbind(psm_treat3, 
                    "Nb_ASV_NA"=psm_treat3_na$Nb_ASV_NA, 
                    "Nb_seq"=psm_treat3_seq$Nb_seq)

psm_treat4 <- psm_treat4 |> 
  mutate("Prop_ASV" = Nb_ASV/(Nb_ASV + Nb_ASV_NA))

ggplot(psm_treat4, aes(y = Function, x = Nb_ASV, fill = Treatment)) +
  geom_bar(stat = "identity", position = "Fill")

ggplot(psm_treat4, aes(y = Function, x = Nb_seq, fill = Treatment)) +
  geom_bar(stat = "identity", position = "Fill")

ggplot(psm_treat4, aes(y = Treatment, x = Prop_ASV, fill = Treatment)) +
  geom_bar(stat = "identity") + facet_wrap(~Function)
```
:::
