---
title: "Bioinformatics pipeline summary"
subtitle: "Where we see the pipeline processes"
date: last-modified
author:
  - name: Adrien Taudière
    url: https://adrientaudiere.github.io/bioinfo.starter/analysis/01_bioinformatics.qmd
    affiliation: IdEst
    affiliation-url: https://orcid.org/my-orcid?orcid=0000-0003-1088-1182
citation: true
# bibliography: bibliography.bib
# link-citations: true
---

# Summary of the bioinformatic pipeline

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(targets)
library(MiscMetabar)
devtools::load_all("~/Nextcloud/IdEst/Projets/MiscMetabar/")
here::i_am("analysis/01_bioinformatics.qmd")
```

<hr/>

The {targets} package is at the core of this project. Please read the intro of the [user manual](https://books.ropensci.org/targets/) if you don't know {targets}.

::: {.cell .page-columns .page-full .column-screen .grey-section}
The {targets} package store ... *targets* in a folder and can load (`tar_load()`) and read (`tar_read`) object from this folder.
:::



```{r}
list.files(tar_read(fastq_files_folder))[!list.files(tar_read(fastq_files_folder))%in%list.files(tar_read(cutadapt))]
```


22MET-01206_S49_L001_R1_001.fastq.gz
22MET-01218_S61_L001_R1_001.fastq.gz 
22MET00155_S5_L001_R1_001.fastq.gz 
22MET00182_S32_L001_R1_001.fastq.gz 
22MET00262_S104_L001_R2_001.fastq.gz 
22MET00268_S110_L001_R2_001.fastq.gz 
22MET00283_S54_L001_R1_001.fastq.gz
22MET00300_S71_L001_R2_001.fastq.gz 
22MET00301_S72_L001_R1_001.fastq.gz 
22MET00353_S83_L001_R2_001.fastq.gz 
22MET01265_S12_L001_R1_001.fastq.gz
22MET01273_S20_L001_R2_001.fastq.gz 
22MET01275_S22_L001_R1_001.fastq.gz
22MET01311_S58_L001_R1_001.fastq.gz 
22MET01336_S83_L001_R1_001.fastq.gz 
22MET01533_S20_L001_R1_001.fastq.gz

22MET1561_S138_L001_R1_001.fastq.gz 
22MET1561_S138_L001_R2_001.fastq.gz 
22MET1562_S139_L001_R1_001.fastq.gz 
22MET1562_S139_L001_R2_001.fastq.gz 
Blanc-22LIBTaschen3_S88_L001_R1_001.fastq.gz 
Blanc-22LIBTaschen3_S88_L001_R2_001.fastq.gz 
Blanc-22LIBTaschen4_S113_L001_R1_001.fastq.gz
Blanc-22LIBTaschen4_S113_L001_R2_001.fastq.gz 
Blanc-Hyphes_S47_L001_R1_001.fastq.gz 
Blanc-Hyphes_S47_L001_R2_001.fastq.gz 
Blanc-Racine_S49_L001_R1_001.fastq.gz
Blanc-Racine_S49_L001_R2_001.fastq.gz




### Load dataset and filter out taxa with less than 10 sequences and present in less than 2 samples 

```{r}
# d_asv (ASV) or d_vs (OTU using vsearch)
d_pq_interm <- tar_read("d_vs", store = here::here("_targets/"))
d_pq_interm2 <- subset_taxa_pq(d_pq_interm, taxa_sums(as_binary_otu_table(d_pq_interm))>1)
d_pq_interm2 <- subset_taxa_pq(d_pq_interm2, taxa_sums(d_pq_interm2)>10)
```


### Filter out samples with less than 2000 sequences

```{r}
barplot(sort(sample_sums(d_pq_interm2)))
barplot(sort(sample_sums(d_pq_interm2))[1:25])
abline(h=2000, col="red")
d_pq_interm2@sam_data$type_ech |> table()
d_pq <- subset_samples_pq(d_pq_interm2, sample_sums(d_pq_interm2)>2000)
d_pq@sam_data$type_ech |> table()
```


```{r}
d_pq <- subset_samples(d_pq, Position!="")
d_pq@sam_data$expe <- "Solace_AM"
```


### Summary of the resulting dataset

```{r}
summary_plot_pq(d_pq)
```

```{r Visualise target plan, out.width='80%'}
tar_glimpse(script = here::here("_targets.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

```{r}
tar_meta(store = here::here("_targets/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```


## Sample data

```{r}
DT::datatable(d_pq@sam_data)
table(d_pq@sam_data$type_ech)
table(d_pq@sam_data$Position, d_pq@sam_data$type_ech)

table(d_pq@sam_data$Bloc)
table(d_pq@sam_data$Tmt, d_pq@sam_data$Type)

table(d_pq@sam_data$G1_G2)
```

## Sequences, samples and clusters across the pipeline


::: {.panel-tabset .nav-pills}
## All samples together

```{r}
kable(tar_read(track_sequences_samples_clusters, store = here::here("_targets/")))
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp <- tar_read(track_by_samples, store = here::here("_targets/"))

for (li in names(tab_samp)) {
  print(knitr::kable(tab_samp[[li]], caption = li, format = "html"))
  cat("\n<!-- -->\n\n")
}
```

:::



### Compare taxonomy using either Marjaam or Eukaryome database

```{r}
krona(d_pq, file = here::here("data", "data_final", "Marjaam.html"), ranks = 2:7, name = "Marjaam")
krona(d_pq, file = here::here("data", "data_final", "Eukaryome.html"), ranks = 10:15, name = "Eukaryome")

merge_krona(c(here::here("data", "data_final", "Marjaam.html"), here::here("data", "data_final", "Eukaryome.html")), here::here("data", "data_final", "compare_tax_database.html"))
```

```{r}
table(d_pq@tax_table[, "Family"])
table(d_pq@tax_table[, "Family_E"])

table(d_pq@tax_table[, "Genus"])
table(d_pq@tax_table[, "Genus_E"])
```


```{r}
psm <- psmelt(d_pq)

ggplot(psm, aes(y = Abundance, x = type_ech, fill = Class_E)) +
  geom_bar(stat = "identity") + no_legend()

plot_tax_pq(
  d_pq,
  "type_ech",
  type = "nb_taxa",
  taxa_fill = "Class_E",
  color_border = rgb(0, 0, 0, 0)
)
```


## Filter AM based on taxonomy

Let's call AMF the Glomeromycota phylum and putative-MFRE the taxa in Densosporales.

```{r}
cond_am_eukaryome_AMF <- d_pq@tax_table[,"Phylum_E"]=="Glomeromycota" 
cond_am_eukaryome_AMF <- as.vector(unclass(cond_am_eukaryome_AMF))
cond_am_eukaryome_AMF[is.na(cond_am_eukaryome_AMF)] <- FALSE
sum(cond_am_eukaryome_AMF)
```


```{r}
sum(d_pq@tax_table[,"Class_E"]=="Endogonomycetes" ,na.rm = T)
sum(d_pq@tax_table[,"Order_E"]=="Densosporales" ,na.rm = T)
sum(d_pq@tax_table[,"Family_E"]=="Endogonaceae" ,na.rm = T)

system(paste0("grep -c 'Endogonaceae' ", 
              here::here("data/data_raw/refseq/DADA2_EUK_SSU_v1.9.fasta")
              ))
system(paste0("grep -c 'Densosporales' ", 
              here::here("data/data_raw/refseq/DADA2_EUK_SSU_v1.9.fasta")
 ))
system(paste0("grep -c 'Endogonomycetes' ", 
              here::here("data/data_raw/refseq/DADA2_EUK_SSU_v1.9.fasta")
 ))
```


```{r}
d_pq2 <- add_new_taxonomy_pq(
  d_pq,
  method = "sintax", 
  ref_fasta = "data/data_raw/refseq/EUK_SSU_v1_9_3.fasta",
  suffix = "_E_sintax")

sum(as.data.frame(as.matrix(d_pq@tax_table))$Order_E_sintax=="Densosporales", na.rm=T)
sum(as.data.frame(as.matrix(d_pq@tax_table))$Family_E_sintax=="Endogonaceae", na.rm=T)

cond_am_eukaryome_mfre <- d_pq@tax_table[,"Family_E"]=="Endogonaceae" | d_pq@tax_table[,"Order_E"]=="Densosporales"
cond_am_eukaryome_mfre <- as.vector(unclass(cond_am_eukaryome_mfre))
cond_am_eukaryome_mfre[is.na(cond_am_eukaryome_mfre)] <- FALSE
sum(cond_am_eukaryome_mfre)
```



```{r}
d_pq <- add_new_taxonomy_pq(
  d_pq2,
  method = "blastn", 
  method_algo = "top-hit",
  ref_fasta = "data/data_raw/refseq/sintax_ref_seq_Mucoromycotina.fasta",
  suffix = "_minimuco_blast_95"
  )

d_pq <- add_new_taxonomy_pq(
  d_pq,
  method = "blastn", 
  min_id = 97,
  method_algo = "top-hit",
  ref_fasta = "data/data_raw/refseq/sintax_ref_seq_Mucoromycotina.fasta",
  suffix = "_minimuco_blast_97"
  )


cond_mfre_miniDB_species <- !is.na(d_pq@tax_table[,"Taxa_name_db_minimuco_blast_95"])
cond_mfre_miniDB_species <- as.vector(unclass(cond_mfre_miniDB_species))
sum(cond_mfre_miniDB_species)

cond_mfre_miniDB_species_97 <- !is.na(d_pq@tax_table[,"Taxa_name_db_minimuco_blast_97"])
cond_mfre_miniDB_species_97 <- as.vector(unclass(cond_mfre_miniDB_species_97))
sum(cond_mfre_miniDB_species_97)

table(cond_mfre_miniDB_species_97, cond_mfre_miniDB_species)

table(cond_am_eukaryome_mfre, cond_mfre_miniDB_species_97)
table(cond_am_eukaryome_AMF, cond_mfre_miniDB_species_97)
table(cond_am_eukaryome_AMF, cond_am_eukaryome_mfre)

ggVennDiagram::ggVennDiagram(list("mfre_euk"=taxa_names(d_pq)[cond_am_eukaryome_mfre],
                                  "mfre_miniDB_97"=taxa_names(d_pq)[cond_mfre_miniDB_species_97],
                                  "mfre_miniDB_95"=taxa_names(d_pq)[cond_mfre_miniDB_species])) +
  scale_x_continuous(expand = expansion(mult = .2))
```


```{r}
# Add an information status of mfre in the tax_table slot
d_pq@tax_table <- tax_table(as.matrix(cbind(d_pq@tax_table, 
                                            "mfre"=cond_am_eukaryome_mfre, 
                                            "amf"=cond_am_eukaryome_AMF)))
```





```{r}
d_am <- clean_pq(subset_taxa(d_pq, amf==TRUE))
d_mfre <- clean_pq(subset_taxa(d_pq, mfre==TRUE))

# Verify that their is no overlap between mfre and amf 
sum(d_am@tax_table[,"mfre"]=="TRUE")
sum(d_mfre@tax_table[,"amf"]=="TRUE")
```


```{r}
table(d_am@tax_table[,"Species_muco_miniDB"])
gtsummary::tbl_summary(dplyr::select(as.data.frame(unclass(d_am@tax_table)), -NA.))

gtsummary::tbl_cross(dplyr::select(as.data.frame(unclass(d_am@tax_table)), c("Genus_E", "Genus")))
gtsummary::tbl_cross(dplyr::select(as.data.frame(unclass(d_mfre@tax_table)), c("Genus_E", "Genus")))
```




## Build a phylogenetic tree

### Using Mafft and fasttree2

#### Only refseq Muco

```{sh}
cd /home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_raw/refseq
mafft --auto --adjustdirection --reorder ref_seq_Mucoromycotina.fasta > ref_seq_Muco_MAFFT.fasta
./FastTreeDbl -gamma -gtr -log fasttree.log ref_seq_Muco_MAFFT.fasta > ref_seq_Muco_MAFFT.tree 
```


```{r}
tree_maft <- ape::read.tree("data/data_raw/refseq/ref_seq_Muco_MAFFT.tree")
plot(tree_maft)
```

#### Outgroup + refseq Muco


```{sh}
cd /home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_raw/refseq
mafft --auto --adjustdirection --reorder ref_seq_Muco_outgroup.fasta > ref_seq_Muco_OG_MAFFT.fasta
./FastTreeDbl -gamma -gtr -log fasttree_OG.log ref_seq_Muco_OG_MAFFT.fasta > ref_seq_Muco_OG_MAFFT.tree 
./raxml-ng_v1.2.2_linux_x86_64/raxml-ng --all --msa ref_seq_Muco_OG_MAFFT.fasta --model GTR+G --tree pars{10} --bs-trees 200 --prefix muco_og_ --bsconverge 
```

```{r}
tree_raxml_muco_og <- read.tree("data/data_raw/refseq/muco_og_.raxml.support")
outgroups <- tree_raxml_muco_og$tip.label[grep("OUT_", tree_raxml_muco_og$tip.label)]
tree_raxml_muco_og <- root(tree_raxml_muco_og, outgroup=outgroups)
plot(tree_raxml_muco_og, show.node.label = TRUE)
```


```{r}
tree_og_maft <- ape::read.tree("data/data_raw/refseq/ref_seq_Muco_OG_MAFFT.tree")
outgroups <- tree_og_maft$tip.label[grep("OUT_", tree_og_maft$tip.label)]
tree_og_maft <- root(tree_og_maft, outgroup=outgroups)
plot(tree_og_maft)
```



#### Outgroup + refseq MFRE + OTU mfre

```{r}
refseq_og <- Biostrings::readDNAStringSet("data/data_raw/refseq/ref_seq_Muco_outgroup.fasta")
seqs_edna_mfre <- Biostrings::DNAStringSet(d_mfre@refseq)

Biostrings::writeXStringSet(c(refseq_og, seqs_edna_mfre), "data/data_intermediate/seqs_mfre_for_mafft.fasta")
Biostrings::writeXStringSet(seqs_edna_mfre, "data/data_intermediate/otu_mfre.fasta")
```


```{sh}
cd /home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_intermediate
mafft --auto --adjustdirection --reorder seqs_mfre_for_mafft.fasta > seqs_mfre_for_mafft_MAFFTED.fasta

/home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_raw/refseq/FastTreeDbl -gamma -gtr -log fasttree_OG.log seqs_mfre_for_mafft_MAFFTED.fasta > seqs_mfre_MAFFTED.tree 

/home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_raw/refseq/raxml-ng_v1.2.2_linux_x86_64/raxml-ng --all --msa seqs_mfre_for_mafft_MAFFTED.fasta --model GTR+G --tree pars{10} --bs-trees 200 --prefix raxml_otu_mfre_og
```


```{r}
library(ape)
tree_raxml_muco_og <- read.tree("data/data_intermediate/raxml_otu_mfre_og.raxml.support")
outgroups <- tree_raxml_muco_og$tip.label[grep("OUT_", tree_raxml_muco_og$tip.label)]
tree_raxml_muco_og <- root(tree_raxml_muco_og, outgroup=outgroups)

nodes_color <- rep("green", length(tree_raxml_muco_og$tip.label))
names(nodes_color) <- tree_raxml_muco_og$tip.label
nodes_color[grepl("Taxa",names(nodes_color))] <- "gray20"
nodes_color[grepl("OUT_",names(nodes_color))] <- "red"

plot(tree_raxml_muco_og, show.node.label = TRUE, tip.color = nodes_color)
```


```{r}
tree_maft <- ape::read.tree("data/data_intermediate/seqs_MAFFTED.tree")
outgroups <- tree_maft$tip.label[grep("OUT_", tree_maft$tip.label)]
tree_maft <- root(tree_maft, outgroup=outgroups)
plot(tree_maft)
```



#### EPA-NG

```{r}
alignment_mafft <-  Biostrings::readDNAStringSet("data/data_intermediate/seqs_mfre_for_mafft_MAFFTED.fasta")
names(alignment_mafft) <- gsub(" .*","",names(alignment_mafft))
Biostrings::writeXStringSet(alignment_mafft[grepl("Taxa",names(alignment_mafft))], "data/data_intermediate/otu_refseq_MAFFTED_with_refseq.fasta")
Biostrings::writeXStringSet(alignment_mafft[!grepl("Taxa",names(alignment_mafft))], "data/data_intermediate/refseq_MAFFTED_with_otu.fasta")
```

```{sh}
cd /home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_raw/refseq

epa-ng --ref-msa /home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_intermediate/refseq_MAFFTED_with_otu.fasta --tree muco_og_.raxml.bestTree --query /home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_intermediate/otu_refseq_MAFFTED_with_refseq.fasta --model muco_og_.raxml.bestModel 
```

```{r}
jplace_tree <- treeio::read.jplace("/home/adrien/Bureau/bioinfo.starter_am_taschen/data/data_raw/refseq/epa_result.jplace")

hist(jplace_tree@placements$like_weight_ratio)
```

```{r}
best_placement <- jplace_tree@placements |> 
  group_by(name) |>
  arrange(desc(like_weight_ratio)) |>
  filter(row_number()==1) 

hist(best_placement$like_weight_ratio)

# outgroups <- c("OUT_X54866.1", "OUT_AY665772.1")
# jplace_tree@phylo <- root(jplace_tree@phylo, outgroup=outgroups)

```

```{r}
library(ggtree)
ggtree(jplace_tree) + geom_tiplab() + ggtree::geom_point(aes(size=ifelse(nplace==0, NA, nplace)))
```

```{r}
epa_res <- jplace_tree@placements |>
  group_by(name) |>
  filter(row_number()==1) |>
 # filter(like_weight_ratio>0.2) |>
  left_join(as_tibble(jplace_tree), by=join_by(node==node)) |> 
  left_join(data.frame(name=names(taxa_sums(d_pq)), n_seq=taxa_sums(d_pq), n_samples=taxa_sums(as_binary_otu_table(d_pq))), by=join_by(name==name))

epa_res |> pull(label) |> table()

epa_res_by_node <- epa_res |>
  group_by(node) |>
  summarise(lwr=mean(like_weight_ratio), taxa=paste(name,collapse = ";"), n_seq=mean(n_seq), n_samples=mean(n_samples), n_otu=n())

jplace_tree2 <- jplace_tree
jplace_tree2@data <- full_join(jplace_tree@data, epa_res_by_node, join_by(node==node))
ggtree(jplace_tree2) + 
  geom_tiplab() + 
  ggtree::geom_point(aes(size=n_otu))
```



```{r}
epa_res2 <- jplace_tree@placements |>
  left_join(as_tibble(jplace_tree), by=join_by(node==node)) |> 
  left_join(data.frame(name=names(taxa_sums(d_pq)), n_seq=taxa_sums(d_pq), n_samples=taxa_sums(as_binary_otu_table(d_pq))), by=join_by(name==name))

epa_res_by_node_2 <- epa_res2 |>
  group_by(node) |>
  summarise(lwr_mean=mean(like_weight_ratio),lwr_sum=sum(like_weight_ratio), taxa=paste(name,collapse = ";"), n_seq=mean(n_seq), n_samples=mean(n_samples), n_otu=n())

jplace_tree3 <- jplace_tree
jplace_tree3@data <- full_join(jplace_tree@data, epa_res_by_node_2, join_by(node==node))
ggtree(jplace_tree3) + 
  geom_tiplab() + 
  ggtree::geom_point(aes(size=lwr_sum, alpha=0.5))
```


```{r}
jplace_tree4 <- jplace_tree
jplace_tree4@data <- left_join(jplace_tree@data, epa_res2, join_by(node==node))

for(taxa_i in taxa_names(d_mfre)) {
p <- ggtree(jplace_tree4) + 
    geom_tiplab(data=jplace_tree4@phylo) + 
    ggtree::geom_point2(aes(size=like_weight_ratio, alpha=0.5, subset = name==taxa_i & !is.na(name)), color="red") +
    labs(title =paste0("Phylogenetic placement for: ", taxa_i, " - (total LWR:" , round(sum(pull(filter(jplace_tree4@data,name==taxa_i & !is.na(name) ), like_weight_ratio)),digits = 3) , ")")) 
print(p)
 }
```


```{r}
jplace_tree@placements |> 
  group_by(name) |>
  summarise(LWR=sum(like_weight_ratio))  |>
  arrange(LWR) |> 
  mutate(mfre_using_euk=name %in% taxa_names(d_pq)[cond_am_eukaryome_mfre],
         mfre_using_blast_95=name %in% taxa_names(d_pq)[cond_mfre_miniDB_species],
         mfre_using_blast_97=name %in% taxa_names(d_pq)[cond_mfre_miniDB_species_97])
  
```


```{r}
venn_data <- list("mfre_euk"=taxa_names(d_pq)[cond_am_eukaryome_mfre],
                                  "mfre_miniDB_97"=taxa_names(d_pq)[cond_mfre_miniDB_species_97],
                                  #"mfre_miniDB_95"=taxa_names(d_pq)[cond_mfre_miniDB_species],
                                  "EPA"=pull(filter(best_placement, like_weight_ratio>0.9), name))
res_venn <- ggVennDiagram::ggVennDiagram(venn_data) +
  scale_x_continuous(expand = expansion(mult = .2))
res_venn

# interactive version
ggVennDiagram::ggVennDiagram(venn_data, names(venn_data), 
              show_intersect = TRUE) 

```




## Trying ASAP (Puillandre et al. 2021) on ASV

```{r}
# load(here("data/data_final/phyloseq_objects.Rdata"))
d_asv <- clean_pq(tar_read("d_asv", store = here::here("_targets/")))
d_asv <- subset_taxa_pq(d_asv, taxa_sums(d_asv)>1000 | taxa_names(d_asv)%in%taxa_names(d_am_mfre))
alignment <- msa::msa(d_asv@refseq, method = "ClustalOmega")

# Save alignement for ASAP software
writeXStringSet(alignment@unmasked, "data/data_intermediate/asv_1000_amf_mfre.fas")

```



```{sh}
cd ~/Bureau/ASAP
./asap ~/Bureau/bioinfo.starter_am_taschen/data/data_intermediate/seqs_for_mafft_MAFFTED.fas -a -o outputs
```












### Using DECIPHER and phangorn

```{r}
seqs_edna <- Biostrings::DNAStringSet(d_mfre@refseq)
refseq_MUCO <- Biostrings::readDNAStringSet("data/data_raw/refseq/ref_seq_Mucoromycotina.fasta")
outgroup <- Biostrings::readDNAStringSet("data/data_raw/refseq/ref_seq_Outgroups.fasta")
names(outgroup) <- paste("OUT_", names(outgroup))

# alignment <-  DECIPHER::AlignSeqs(c(seqs_edna, refseq_MUCO, outgroup), anchor = NA)
# phang.align <- phangorn::phyDat(as(alignment, "matrix"), type = "DNA")

alignment <- msa::msa(c(seqs_edna, refseq_MUCO, outgroup), method = "ClustalW")@unmasked
phang.align <- phangorn::phyDat(as(alignment, "matrix"), type = "DNA")

```


```{r}
dm <- phangorn::dist.ml(phang.align)
treeNJ <- phangorn::NJ(dm) # Note, tip order != sequence order
outgroups <- treeNJ$tip.label[grep("OUT_", treeNJ$tip.label)]
treeNJ <- root(treeNJ, outgroup=outgroups)

fit <- phangorn::pml(treeNJ, data = phang.align)

fitGTR <- update(fit, k = 4, inv = 0.2)
tree_ML <-
  phangorn::optim.pml(
    fitGTR,
    model = "GTR",
    optInv = TRUE,
    optGamma = TRUE,
    rearrangement = "NNI",
    control = phangorn::pml.control(trace = 0),
    optNni = TRUE
  )

tree_ML_bs <- phangorn::bootstrap.pml(
        tree_ML,
        bs = 100,
        multicore = TRUE,
        rearrangement = "NNI"
      )

```


```{r}
nodes_color <- rep(NA, length(alignment))
names(nodes_color) <- names(alignment)
nodes_color[names(nodes_color) %in% names(seqs_edna)] <- "gray20"
nodes_color[names(nodes_color) %in% names(refseq_MUCO)] <- "green"
nodes_color[names(nodes_color) %in% names(outgroup)] <- "red"

phangorn::plotBS(tree_ML, tree_ML_bs, "phylogram", tip.color=nodes_color)
phangorn::densiTree(tree_ML_bs, type="cladogram", col="blue", alpha = 0.1)
```



```{r}
d_am_mfre <- clean_pq(subset_taxa(d_pq, amf==TRUE | mfre==TRUE))
track_wkflow(list("All sequences" = d_pq, "AMF_MFRE" = d_am_mfre, "AMF" = clean_pq(d_am), "MFRE"=clean_pq(d_mfre)))
```

```{r}
save(d_pq, d_am_mfre, d_am, d_mfre, file=here::here("data/data_final/phyloseq_objects.Rdata"))
```



```{r}
l_pq <- list("OTU vs"=d_pq_interm, 
             "OTU with at least 11 sequences and 2 occurences" = d_pq_interm2,
             "Samples with more than 2000 sequences" = clean_pq(d_pq),
             "AM"= clean_pq(d_am), 
             "MFRE"=clean_pq(d_mfre), 
             "AM and MFRE"=clean_pq(d_am_mfre))

cbind(rbind(tar_read(track_sequences_samples_clusters), 
      "---" =c("---", "---", "---"),
      track_wkflow(l_pq)), 
      "Sol samples" = unlist(c(rep(NA, 7), "---", as.numeric(lapply(l_pq, function(x) {sum(x@sam_data$type_ech=="sol")}) ))), 
      "Hyphes samples" = unlist(c(rep(NA, 7), "---", as.numeric(lapply(l_pq, function(x) {sum(x@sam_data$type_ech=="hyphes")}) ))),
      "Root samples"= unlist(c(rep(NA, 7), "---", as.numeric(lapply(l_pq, function(x) {sum(x@sam_data$type_ech=="ADN_Racines")}) ))))
```



{{< include _session_info.qmd >}}
