---
title: "Bioinformatics pipeline summary"
subtitle: "Where we see the pipeline processes"
date: last-modified
author:
  - name: Adrien Taudière
    url: https://adrientaudiere.github.io/bioinfo.starter/analysis/01_bioinformatics.qmd
    affiliation: IdEst
    affiliation-url: https://orcid.org/my-orcid?orcid=0000-0003-1088-1182
citation: true
# bibliography: bibliography.bib
# link-citations: true
---

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(targets)
library(MiscMetabar)
here::i_am("analysis/01_bioinformatics.qmd")
source(here::here("R/functions.R"))
```

```{r child="../R/styles.qmd"}

```

# Archaea 16S

Marker: Archaea 16S rRNA gene (primers Arch349F / Arch806R). Targets store: `_targets/`.

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga <- greenAlgoR::ga_targets(store = here::here("_targets/"))
print(paste0(res_ga$carbon_footprint_total_gCO2, " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data <- autometric::log_read(
  here::here("data/data_final/autometric_log.txt"),
  units_time = "hours",
  units_memory = "gigabytes"
)

log_data |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  ggplot(aes(x = time, y = reorder(phase, desc(time)), color = resident)) +
  geom_line(aes(linewidth = cpu)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  theme_idest() +
  xlab("Time (in hours)") +
  ylab("")
```

## Targets plan

```{r Visualise target plan archaea, out.width='80%'}
tar_glimpse(script = here::here("_targets.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

## Phyloseq object

```{r}
d_pq_archaea <- clean_pq(tar_read("d_vs_mumu", store = here::here("_targets/")))
```

```{r}
summary_plot_pq(d_pq_archaea)
```

## Targets metadata

```{r}
tar_meta(store = here::here("_targets/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```

## Sample data

```{r}
DT::datatable(d_pq_archaea@sam_data)
```

## Sequences, samples and clusters across the pipeline

::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_16S <- tar_read(track_sequences_samples_clusters, store = here::here("_targets/"))
kable(track_16S)
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp <- tar_read(track_by_samples, store = here::here("_targets/"))

for (li in names(tab_samp)) {
  print(knitr::kable(tab_samp[[li]], caption = li, format = "html"))
  cat('\n<!-- -->\n\n')
}
```

:::

## Krona charts

```{r}
krona(
  clean_pq(d_pq_archaea, simplify_taxo = TRUE),
  file = here::here("data/data_final/krona_archaea_nb_seq.html")
)
krona(
  clean_pq(d_pq_archaea, simplify_taxo = TRUE),
  nb_seq = FALSE,
  file = here::here("data/data_final/krona_archaea_nb_taxa.html")
)
```

<hr/>

# mcrA — Methane production

Marker: mcrA gene (methyl-coenzyme M reductase, primers GGTGGTGTMGGDTTCACMCARTA / CGTTCATBGCGTAGTTVGGRTAGT). No taxonomic assignment for this marker. Targets store: `_targets_mcrA/`.

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga_mcrA <- greenAlgoR::ga_targets(store = here::here("_targets_mcrA/"))
print(paste0(res_ga_mcrA$carbon_footprint_total_gCO2, " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data_mcrA <- autometric::log_read(
  here::here("data/data_final/autometric_log_mcrA.txt"),
  units_time = "hours",
  units_memory = "gigabytes"
)

log_data_mcrA |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  ggplot(aes(x = time, y = reorder(phase, desc(time)), color = resident)) +
  geom_line(aes(linewidth = cpu)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  theme_idest() +
  xlab("Time (in hours)") +
  ylab("")
```

## Targets plan

```{r Visualise target plan mcrA, out.width='80%'}
tar_glimpse(script = here::here("_targets_mcrA.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

## Phyloseq object

```{r}
d_pq_mcrA <- clean_pq(tar_read("d_vs", store = here::here("_targets_mcrA/")))
```

```{r}
summary_plot_pq(d_pq_mcrA)
```

## Targets metadata

```{r}
tar_meta(store = here::here("_targets_mcrA/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```

## Sample data

```{r}
DT::datatable(d_pq_mcrA@sam_data)
```

## Sequences, samples and clusters across the pipeline

::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_mcrA <- tar_read(track_sequences_samples_clusters, store = here::here("_targets_mcrA/"))
kable(track_mcrA)
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp_mcrA <- tar_read(track_by_samples, store = here::here("_targets_mcrA/"))

for (li in names(tab_samp_mcrA)) {
  print(knitr::kable(tab_samp_mcrA[[li]], caption = li, format = "html"))
  cat('\n<!-- -->\n\n')
}
```

:::

<hr/>

# pmoA — Methane oxidation

Marker: pmoA gene (particulate methane monooxygenase, primers GGNGACTGGGACTTCTGG / CCGGMGCAACGTCYTTACC). No taxonomic assignment for this marker. Targets store: `_targets_pmoA/`.

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga_pmoA <- greenAlgoR::ga_targets(store = here::here("_targets_pmoA/"))
print(paste0(res_ga_pmoA$carbon_footprint_total_gCO2, " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data_pmoA <- autometric::log_read(
  here::here("data/data_final/autometric_log_pmoA.txt"),
  units_time = "hours",
  units_memory = "gigabytes"
)

log_data_pmoA |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  ggplot(aes(x = time, y = reorder(phase, desc(time)), color = resident)) +
  geom_line(aes(linewidth = cpu)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  theme_idest() +
  xlab("Time (in hours)") +
  ylab("")
```

## Targets plan

```{r Visualise target plan pmoA, out.width='80%'}
tar_glimpse(script = here::here("_targets_pmoA.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

## Phyloseq object

```{r}
d_pq_pmoA <- clean_pq(tar_read("d_vs", store = here::here("_targets_pmoA/")))
```

```{r}
summary_plot_pq(d_pq_pmoA)
```

## Targets metadata

```{r}
tar_meta(store = here::here("_targets_pmoA/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```

## Sample data

```{r}
DT::datatable(d_pq_pmoA@sam_data)
```

## Sequences, samples and clusters across the pipeline

::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_pmoA <- tar_read(track_sequences_samples_clusters, store = here::here("_targets_pmoA/"))
kable(track_pmoA)
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp_pmoA <- tar_read(track_by_samples, store = here::here("_targets_pmoA/"))

for (li in names(tab_samp_pmoA)) {
  print(knitr::kable(tab_samp_pmoA[[li]], caption = li, format = "html"))
  cat('\n<!-- -->\n\n')
}
```

:::

<hr/>

# Cross-marker comparison

Summary of read and cluster counts across the three markers at the final pipeline step (OTU vsearch + mumu).

## Raw nb of sequences

```{r}
df_raw_seq <- t(tibble("Total"= track_16S[1,"nb_sequences"],
        "16S"= track_16S[2,"nb_sequences"],
        "mcrA"= track_mcrA[2,"nb_sequences"],
        "pmoA"= track_pmoA[2,"nb_sequences"],
        "Unknown"= track_16S[1,"nb_sequences"] - track_16S[2,"nb_sequences"] - track_mcrA[2,"nb_sequences"] - track_pmoA[2,"nb_sequences"]
        ) ) |>
          mutate("%_total"= round(Total/Total*100, 2),
                 "%_16S"= round(16S/Total*100, 2),
                 "%_mcrA"= round(mcrA/Total*100, 2),
                 "%_pmoA"= round(pmoA/Total*100, 2),
                 "%_Unknown"= round(Unknown/Total*100, 2)
                 ) 

kable(df_raw_seq, caption = "Number of sequences after cutadapt filtering step for each marker")
```


## Sequences and clusters

```{r}
track_wkflow(list("16S"=d_16S,
                  "pmoA"=d_pq_pmoA, 
                  "mcrA"=d_pq_mcrA
                  ))
```

## OTU richness per sample

```{r}
richness_archaea <- data.frame(
  Marker = "Archaea 16S",
  Sample = sample_names(d_pq_archaea),
  OTUs = estimate_richness(d_pq_archaea, measures = "Observed")$Observed
)
richness_mcrA <- data.frame(
  Marker = "mcrA",
  Sample = sample_names(d_pq_mcrA),
  OTUs = estimate_richness(d_pq_mcrA, measures = "Observed")$Observed
)
richness_pmoA <- data.frame(
  Marker = "pmoA",
  Sample = sample_names(d_pq_pmoA),
  OTUs = estimate_richness(d_pq_pmoA, measures = "Observed")$Observed
)

richness_all <- rbind(richness_archaea, richness_mcrA, richness_pmoA)

ggplot(richness_all, aes(x = Marker, y = OTUs, fill = Marker)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 1.5, alpha = 0.6) +
  theme_idest() +
  xlab("") +
  ylab("OTU richness per sample") +
  theme(legend.position = "none")
```

## Read counts per sample

```{r}
reads_archaea <- data.frame(
  Marker = "Archaea 16S",
  Sample = sample_names(d_pq_archaea),
  Reads = sample_sums(d_pq_archaea)
)
reads_mcrA <- data.frame(
  Marker = "mcrA",
  Sample = sample_names(d_pq_mcrA),
  Reads = sample_sums(d_pq_mcrA)
)
reads_pmoA <- data.frame(
  Marker = "pmoA",
  Sample = sample_names(d_pq_pmoA),
  Reads = sample_sums(d_pq_pmoA)
)

reads_all <- rbind(reads_archaea, reads_mcrA, reads_pmoA)

ggplot(reads_all, aes(x = Marker, y = Reads, fill = Marker)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 1.5, alpha = 0.6) +
  theme_idest() +
  xlab("") +
  ylab("Read count per sample") +
  theme(legend.position = "none")
```

## Carbon footprint summary

```{r}
carbon_table <- data.frame(
  Marker = c("Archaea 16S", "mcrA", "pmoA"),
  gCO2_equiv = c(
    res_ga$carbon_footprint_total_gCO2,
    res_ga_mcrA$carbon_footprint_total_gCO2,
    res_ga_pmoA$carbon_footprint_total_gCO2
  )
)
kable(carbon_table, col.names = c("Marker", "Carbon footprint (gCO2 equiv.)"))
```

{{< include _session_info.qmd >}}
