---
title: "Bioinformatics pipeline summary"
subtitle: "Where we see the pipeline processes"
date: last-modified
author:
  - name: Adrien Taudière
    url: https://adrientaudiere.github.io/bioinfo.starter/analysis/01_bioinformatics.qmd
    affiliation: IdEst
    affiliation-url: https://orcid.org/my-orcid?orcid=0000-0003-1088-1182
citation: true
# bibliography: bibliography.bib
# link-citations: true
---

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(targets)
library(MiscMetabar)
library(patchwork)

here::i_am("analysis/01_bioinformatics.qmd")
source(here::here("R/functions.R"))
source(here::here("R/styles.R"))
```

```{r child="../R/styles.qmd"}

```

# Archaea 16S

Marker: Archaea 16S rRNA gene (primers Arch349F / Arch806R). Targets store: `_targets/`.

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga <- greenAlgoR::ga_targets(store = here::here("_targets/"), PSF=2)
print(paste0(round(res_ga$carbon_footprint_total_gCO2, 0), " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data <- autometric::log_read(
  here::here("data/data_final/autometric_log.txt"),
  units_time = "hours",
  units_memory = "gigabytes"
)

log_data |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  ggplot(aes(x = time, y = reorder(phase, desc(time)), color = resident)) +
  geom_line(aes(linewidth = cpu)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  theme_idest() +
  xlab("Time (in hours)") +
  ylab("")
```

## Targets plan

```{r Visualise target plan archaea, out.width='80%'}
tar_glimpse(script = here::here("_targets.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

## Phyloseq object

```{r}
d_16S <- clean_pq(tar_read("d_vs_mumu", store = here::here("_targets/")))
```

```{r}
summary_plot_pq(d_16S)
```

## Targets metadata

```{r}
tar_meta(store = here::here("_targets/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```

## Sample data

```{r}
DT::datatable(d_16S@sam_data)
```


## Denovo chimera removal (second time !)

```{r}
d_16S <- MiscMetabar::chimera_removal_vs(d_16S)
```

## Control samples

```{r}
d_16S@sam_data$is_neg_control <- ifelse(grepl("Control", d_16S@sam_data$Numeric_code), TRUE, FALSE)

tidypq::neg_control_diag_pq(d_16S, is_neg_control)

d_16S_wo_neg_control <- tidypq::neg_control_clean_pq(d_16S, is_neg_control)
```


## Sequences, samples and clusters across the pipeline

::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_16S <- tar_read(track_sequences_samples_clusters, store = here::here("_targets/"))
kable(track_16S)
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp <- tar_read(track_by_samples, store = here::here("_targets/"))

for (li in names(tab_samp)) {
  print(knitr::kable(tab_samp[[li]], caption = li, format = "html"))
  cat('\n<!-- -->\n\n')
}
```

:::

## Krona charts

```{r}
krona(
  clean_pq(d_16S, simplify_taxo = TRUE),
  file = here::here("data/data_final/krona_archaea_nb_seq.html")
)
krona(
  clean_pq(d_16S, simplify_taxo = TRUE),
  nb_seq = FALSE,
  file = here::here("data/data_final/krona_archaea_nb_taxa.html")
)
```

<hr/>

# mcrA — Methane production

Marker: mcrA gene (methyl-coenzyme M reductase, primers GGTGGTGTMGGDTTCACMCARTA / CGTTCATBGCGTAGTTVGGRTAGT). No taxonomic assignment for this marker. Targets store: `_targets_mcrA/`.

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga_mcrA <- greenAlgoR::ga_targets(store = here::here("_targets_mcrA/"), PSF=2)
print(paste0(round(res_ga_mcrA$carbon_footprint_total_gCO2, 0), " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data_mcrA <- autometric::log_read(
  here::here("data/data_final/autometric_log_mcrA.txt"),
  units_time = "hours",
  units_memory = "gigabytes"
)

log_data_mcrA |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  ggplot(aes(x = time, y = reorder(phase, desc(time)), color = resident)) +
  geom_line(aes(linewidth = cpu)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  theme_idest() +
  xlab("Time (in hours)") +
  ylab("")
```

## Targets plan

```{r Visualise target plan mcrA, out.width='80%'}
tar_glimpse(script = here::here("_targets_mcrA.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

## Phyloseq object

```{r}
d_mcrA <- clean_pq(tar_read("d_vs", store = here::here("_targets_mcrA/")))
```

```{r}
summary_plot_pq(d_mcrA)
```

## Targets metadata

```{r}
tar_meta(store = here::here("_targets_mcrA/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```

## Sample data

```{r}
DT::datatable(d_mcrA@sam_data)
```

## Denovo chimera removal (second time !)

```{r}
d_mcrA <- MiscMetabar::chimera_removal_vs(d_mcrA)
```

## Control samples

```{r}
d_mcrA@sam_data$is_neg_control <- ifelse(grepl("Control", d_mcrA@sam_data$Numeric_code), TRUE, FALSE)

tidypq::neg_control_diag_pq(d_mcrA, is_neg_control)

d_mcrA_wo_neg_control <- tidypq::neg_control_clean_pq(d_mcrA, is_neg_control)
```


## Sequences, samples and clusters across the pipeline

::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_mcrA <- tar_read(track_sequences_samples_clusters, store = here::here("_targets_mcrA/"))
kable(track_mcrA)
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp_mcrA <- tar_read(track_by_samples, store = here::here("_targets_mcrA/"))

for (li in names(tab_samp_mcrA)) {
  print(knitr::kable(tab_samp_mcrA[[li]], caption = li, format = "html"))
  cat('\n<!-- -->\n\n')
}
```

:::

<hr/>

# pmoA — Methane oxidation

Marker: pmoA gene (particulate methane monooxygenase, primers GGNGACTGGGACTTCTGG / CCGGMGCAACGTCYTTACC). No taxonomic assignment for this marker. Targets store: `_targets_pmoA/`.

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga_pmoA <- greenAlgoR::ga_targets(store = here::here("_targets_pmoA/"), PSF=2)
print(paste0(round(res_ga_pmoA$carbon_footprint_total_gCO2, 0), " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data_pmoA <- autometric::log_read(
  here::here("data/data_final/autometric_log_pmoA.txt"),
  units_time = "hours",
  units_memory = "gigabytes"
)

log_data_pmoA |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  ggplot(aes(x = time, y = reorder(phase, desc(time)), color = resident)) +
  geom_line(aes(linewidth = cpu)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  theme_idest() +
  xlab("Time (in hours)") +
  ylab("")
```

## Targets plan

```{r Visualise target plan pmoA, out.width='80%'}
tar_glimpse(script = here::here("_targets_pmoA.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

## Phyloseq object

```{r}
d_pmoA <- clean_pq(tar_read("d_vs", store = here::here("_targets_pmoA/")))
```

```{r}
summary_plot_pq(d_pmoA)
```

## Targets metadata

```{r}
tar_meta(store = here::here("_targets_pmoA/"), targets_only = TRUE) |>
  dplyr::mutate(time = paste0(seconds %/% 3600, ":", seconds %/% 60, ":", floor(seconds %% 60))) |>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb = round(bytes / 10^9, 2)) |>
  dplyr::arrange(desc(seconds), desc(bytes)) |>
  kable()
```

## Sample data

```{r}
DT::datatable(d_pmoA@sam_data)
```


## Denovo chimera removal (second time !)

```{r}
d_pmoA <- MiscMetabar::chimera_removal_vs(d_pmoA)
```

## Control samples

```{r}
d_pmoA@sam_data$is_neg_control <- ifelse(grepl("Control", d_pmoA@sam_data$Numeric_code), TRUE, FALSE)

tidypq::neg_control_diag_pq(d_pmoA, is_neg_control)

d_pmoA_wo_neg_control <- tidypq::neg_control_clean_pq(d_pmoA, is_neg_control)
```

## Sequences, samples and clusters across the pipeline

::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_pmoA <- tar_read(track_sequences_samples_clusters, store = here::here("_targets_pmoA/"))
kable(track_pmoA)
```

## Per samples

```{r}
#| message: false
#| results: asis
tab_samp_pmoA <- tar_read(track_by_samples, store = here::here("_targets_pmoA/"))

for (li in names(tab_samp_pmoA)) {
  print(knitr::kable(tab_samp_pmoA[[li]], caption = li, format = "html"))
  cat('\n<!-- -->\n\n')
}
```

:::

<hr/>

# Cross-marker comparison

Summary of read and cluster counts across the three markers at the final pipeline step (OTU vsearch + mumu).

## Raw nb of sequences

```{r}
df_raw_seq <- tibble(data.frame("Total"= track_16S[1,"nb_sequences"],
        "mark_16S"= track_16S[2,"nb_sequences"],
        "mcrA"= track_mcrA[2,"nb_sequences"],
        "pmoA"= track_pmoA[2,"nb_sequences"],
        "Unknown"= track_16S[1,"nb_sequences"] - track_16S[2,"nb_sequences"] - track_mcrA[2,"nb_sequences"] - track_pmoA[2,"nb_sequences"]
        )) |>                                                                
    tidyr::pivot_longer(everything(), names_to = "variable", values_to = "Number") |>       
          mutate("Percent" = paste0(round(100*Number/Number[1], 2), "%"))

kable(df_raw_seq, caption = "Number of sequences after cutadapt filtering step for each marker. Negative percent in Unknown indicate redundant sequences")
```

## Verify for the presence or other markers in each refseq

| Project | Script | Marker | Primers | Store |
|---------|--------|--------|---------|-------|
| `main`  | `_targets.R` | Archaea 16S (Arch349F / Arch806R) | `CCCTACGGGGTGCASCAG` / `GGACTACVSGGGTATCTAAT` | `_targets/` |
| `mcrA`  | `_targets_mcrA.R` | mcrA — methane production | `GGTGGTGTMGGDTTCACMCARTA` / `CGTTCATBGCGTAGTTVGGRTAGT` | `_targets_mcrA/` |
| `pmoA`  | `_targets_pmoA.R` | pmoA — methane oxidation | `GGNGACTGGGACTTCTGG` / `CCGGMGCAACGTCYTTACC` | `_targets_pmoA/` |

```{r}
#| cache: true
primers_cross <- c(
  mcrA_fwd = "GGTGGTGTMGGDTTCACMCARTA",
  mcrA_rev = "CGTTCATBGCGTAGTTVGGRTAGT",
  pmoA_fwd = "GGNGACTGGGACTTCTGG",
  pmoA_rev = "CCGGMGCAACGTCYTTACC"
)

check_primers_in_refseq <- function(primers, ps_list) {
  do.call(rbind, lapply(names(ps_list), function(ps_name) {
    refseq <- ps_list[[ps_name]]@refseq
    do.call(rbind, lapply(names(primers), function(p_name) {
      p <- Biostrings::DNAString(primers[[p_name]])
      rc_p <- Biostrings::reverseComplement(p)
      fwd <- Biostrings::vcountPattern(p, refseq, fixed = FALSE) > 0
      rev <- Biostrings::vcountPattern(rc_p, refseq, fixed = FALSE) > 0
      data.frame(
        phyloseq = ps_name,
        primer = p_name,
        sequence = primers[[p_name]],
        n_fwd = sum(fwd),
        n_rev = sum(rev),
        n_any = sum(fwd | rev)
      )
    }))
  }))
}

d_16S <- clean_pq(tar_read("d_vs", store = here::here("_targets/")))

primer_check <- check_primers_in_refseq(
  primers_cross,
  list("16S" = d_16S, "pmoA" = d_pmoA, "mcrA" = d_mcrA)
)
kable(primer_check, caption = "Number of refseq sequences containing each cross-marker primer (fwd or rc). fixed = FALSE expands IUPAC codes (M, R, Y, …).")
```

```{r}
#| cache: true
all_primers <- c(
  archaea_16S_fwd = "CCCTACGGGGTGCASCAG",
  archaea_16S_rev = "GGACTACVSGGGTATCTAAT",
  mcrA_fwd        = "GGTGGTGTMGGDTTCACMCARTA",
  mcrA_rev        = "CGTTCATBGCGTAGTTVGGRTAGT",
  pmoA_fwd        = "GGNGACTGGGACTTCTGG",
  pmoA_rev        = "CCGGMGCAACGTCYTTACC"
)

#' Find taxa whose refseq contains any primer sequence
#'
#' For each taxon in `ps`, checks whether the representative sequence in
#' `@refseq` contains any primer from `primers` (forward or reverse complement),
#' using IUPAC-aware pattern matching (`fixed = FALSE`).
#'
#' @param ps A phyloseq object with a populated `@refseq` slot.
#' @param primers Named character vector of primer sequences (IUPAC codes
#'   allowed, e.g. M, R, Y, N).
#'
#' @return A data.frame with columns `taxon`, `matched_primers` (comma-separated
#'   names of all primers that matched), and `n_reads` (total reads across
#'   samples). Returns `NULL` if no contaminated taxa are found.
find_taxa_with_primers <- function(ps, primers) {
  refseq <- ps@refseq
  match_mat <- vapply(names(primers), function(p_name) {
    p <- Biostrings::DNAString(primers[[p_name]])
    rc_p <- Biostrings::reverseComplement(p)
    fwd <- Biostrings::vcountPattern(p, refseq, fixed = FALSE) > 0
    rev <- Biostrings::vcountPattern(rc_p, refseq, fixed = FALSE) > 0
    fwd | rev
  }, logical(length(refseq)))

  any_match <- rowSums(match_mat) > 0
  if (!any(any_match)) {
    return(NULL)
  }
  bad_idx <- which(any_match)
  data.frame(
    taxon = names(refseq)[bad_idx],
    matched_primers = apply(match_mat[bad_idx, , drop = FALSE], 1, \(row) paste(names(primers)[row], collapse = ", ")),
    n_reads = phyloseq::taxa_sums(ps)[bad_idx],
    row.names = NULL
  )
}

bad_taxa <- list(
  "16S"  = find_taxa_with_primers(d_16S, all_primers),
  "pmoA" = find_taxa_with_primers(d_pmoA, all_primers),
  "mcrA" = find_taxa_with_primers(d_mcrA, all_primers)
)

for (ps_name in names(bad_taxa)) {
  df <- bad_taxa[[ps_name]]
  if (is.null(df) || nrow(df) == 0) {
    message(ps_name, ": no taxa with primer contamination.")
  } else {
    print(kable(
      df[order(-df$n_reads), ],
      caption = sprintf("%s — %d taxa containing a primer sequence (%d total reads)", ps_name, nrow(df), sum(df$n_reads))
    ))
  }
}
```

To filter out the contaminated taxa, use `prune_taxa()`:

```{r}
filter_primer_taxa <- function(ps, primers) {
  bad <- find_taxa_with_primers(ps, primers)
  if (is.null(bad) || nrow(bad) == 0) {
    return(ps)
  }
  message("Removing ", nrow(bad), " taxa containing primer sequences (", sum(bad$n_reads), " reads).")
  phyloseq::prune_taxa(!phyloseq::taxa_names(ps) %in% bad$taxon, ps)
}

d_16S  <- filter_primer_taxa(d_16S,  all_primers)
d_pmoA <- filter_primer_taxa(d_pmoA, all_primers)
d_mcrA <- filter_primer_taxa(d_mcrA, all_primers)
```

```{r}
#| cache: true

#' Compare refseq sequences across three phyloseq objects
#'
#' Computes pairwise sequence similarity between the `@refseq` slots of three
#' phyloseq objects using k-mer Jaccard similarity, then groups near-identical
#' sequences via connected components (igraph). Produces a ggplot2 line plot
#' and a Venn diagram (ggVennDiagram) for each dissimilarity threshold.
#'
#' Similarity is approximated as `p ≈ (2j / (1 + j))^(1/k)` where `j` is the
#' Jaccard index of shared k-mer presence and `k` is the k-mer size. The
#' approximation is accurate within ~2 pp for identities between 90–100%.
#'
#' @param ps1,ps2,ps3 phyloseq objects whose `@refseq` slot will be compared.
#' @param labels Character vector of length 3. Display names for `ps1`, `ps2`,
#'   and `ps3` respectively. Default: `c("ps1", "ps2", "ps3")`.
#' @param thresholds Numeric vector of maximum dissimilarity thresholds (in %).
#'   `0` means exact matches only; `10` means sequences within 10% divergence
#'   are considered equivalent. Default: `c(0, 1, 2, 3, 5, 10)`.
#' @param k Integer. K-mer size used to build presence/absence profiles.
#'   Larger `k` increases specificity; smaller `k` increases sensitivity for
#'   divergent sequences. Default: `7`.
#'
#' @return A named list with three elements:
#'   \describe{
#'     \item{`line_plot`}{ggplot2 line plot: x = number of shared sequence
#'       clusters, y = minimum similarity threshold, one line per pairwise
#'       comparison.}
#'     \item{`venn_plots`}{Named list of ggVennDiagram objects, one per
#'       threshold. Names follow the pattern `"<threshold>_pct"`.}
#'     \item{`results`}{List of per-threshold results, each containing
#'       `threshold`, `venn_data`, `n12`, `n13`, and `n23` (number of shared
#'       sequence clusters between each pair).}
#'   }
compare_refseq_similarity <- function(ps1, ps2, ps3,
                                      labels = c("ps1", "ps2", "ps3"),
                                      thresholds = c(0, 1, 2, 3, 5, 10),
                                      k = 7) {
  seq1 <- as.character(ps1@refseq)
  seq2 <- as.character(ps2@refseq)
  seq3 <- as.character(ps3@refseq)
  n1 <- length(seq1)
  n2 <- length(seq2)
  n3 <- length(seq3)

  kmer_presence <- function(seqs) {
    dna <- Biostrings::DNAStringSet(seqs)
    (Biostrings::oligonucleotideFrequency(dna, width = k) > 0) * 1
  }

  jaccard_to_pid <- function(jaccard) {
    100 * (2 * jaccard / (1 + jaccard))^(1 / k)
  }

  compute_pid_matrix <- function(km_a, km_b, n_a, n_b) {
    shared <- tcrossprod(km_a, km_b)
    union_size <- outer(n_a, n_b, "+") - shared
    jaccard_to_pid(shared / pmax(union_size, 1))
  }

  message("Building k-mer profiles (k = ", k, ")...")
  km1 <- kmer_presence(seq1)
  km2 <- kmer_presence(seq2)
  km3 <- kmer_presence(seq3)
  n1k <- rowSums(km1)
  n2k <- rowSums(km2)
  n3k <- rowSums(km3)

  message("Computing ", labels[1], " vs ", labels[2], "...")
  pid12 <- compute_pid_matrix(km1, km2, n1k, n2k)
  message("Computing ", labels[1], " vs ", labels[3], "...")
  pid13 <- compute_pid_matrix(km1, km3, n1k, n3k)
  message("Computing ", labels[2], " vs ", labels[3], "...")
  pid23 <- compute_pid_matrix(km2, km3, n2k, n3k)

  process_threshold <- function(thresh) {
    if (thresh == 0) {
      all_seqs <- c(seq1, seq2, seq3)
      seq_groups <- split(seq_along(all_seqs), all_seqs)
      edges <- do.call(rbind, lapply(seq_groups, function(idx) {
        if (length(idx) < 2) {
          return(NULL)
        }
        t(combn(idx, 2))
      }))
    } else {
      min_pid_val <- 100 - thresh
      e12 <- which(pid12 >= min_pid_val, arr.ind = TRUE)
      e13 <- which(pid13 >= min_pid_val, arr.ind = TRUE)
      e23 <- which(pid23 >= min_pid_val, arr.ind = TRUE)
      if (nrow(e12) > 0) {
        e12[, 2] <- e12[, 2] + n1
      }
      if (nrow(e13) > 0) {
        e13[, 2] <- e13[, 2] + n1 + n2
      }
      if (nrow(e23) > 0) {
        e23[, 1] <- e23[, 1] + n1
        e23[, 2] <- e23[, 2] + n1 + n2
      }
      edges <- rbind(e12, e13, e23)
    }

    g <- igraph::make_empty_graph(n = n1 + n2 + n3, directed = FALSE)
    if (!is.null(edges) && nrow(edges) > 0) {
      g <- igraph::add_edges(g, t(edges))
    }
    comp_id <- igraph::components(g)$membership

    node_set <- rep(labels, c(n1, n2, n3))
    ct <- table(comp_id, node_set)
    in_A <- ct[, labels[1]] > 0
    in_B <- ct[, labels[2]] > 0
    in_C <- ct[, labels[3]] > 0
    comp_names <- paste0("c", rownames(ct))

    list(
      threshold = thresh,
      venn_data = stats::setNames(
        list(comp_names[in_A], comp_names[in_B], comp_names[in_C]),
        labels
      ),
      n12 = sum(in_A & in_B),
      n13 = sum(in_A & in_C),
      n23 = sum(in_B & in_C)
    )
  }

  results <- lapply(thresholds, process_threshold)

  df_line <- do.call(rbind, lapply(results, function(r) {
    data.frame(
      similarity = 100 - r$threshold,
      n_shared = c(r$n12, r$n13, r$n23),
      comparison = c(
        paste(labels[1], "vs", labels[2]),
        paste(labels[1], "vs", labels[3]),
        paste(labels[2], "vs", labels[3])
      )
    )
  }))

  p_line <- ggplot2::ggplot(
    df_line,
    ggplot2::aes(x = n_shared, y = similarity, color = comparison, group = comparison)
  ) +
    ggplot2::geom_line(linewidth = 1.2) +
    ggplot2::geom_point(size = 3) +
    ggplot2::scale_y_continuous(
      breaks = 100 - thresholds,
      labels = paste0(100 - thresholds, "%")
    ) +
    ggplot2::labs(
      x = "Number of shared sequence clusters",
      y = "Minimum sequence similarity",
      color = "Comparison"
    )

  venn_plots <- lapply(results, function(r) {
    thresh <- r$threshold
    ggVennDiagram::ggVennDiagram(r$venn_data, label_alpha = 0, label = "count") +
      ggplot2::scale_fill_gradient(low = "white", high = "#4CAF50") +
      ggplot2::labs(
        title = if (thresh == 0) "Exact matches" else sprintf("Within %d%% dissim.", thresh)
      ) +
      ggplot2::theme(legend.position = "none")
  })
  names(venn_plots) <- paste0(thresholds, "_pct")

  list(line_plot = p_line, venn_plots = venn_plots, results = results)
}

sim_results <- compare_refseq_similarity(
  d_16S, d_pmoA, d_mcrA,
  labels = c("16S", "pmoA", "mcrA"),
  thresholds = c(0)
)
```

```{r}
#| fig-width: 14
#| fig-height: 9
patchwork::wrap_plots(sim_results$venn_plots, ncol = 3)
```

## Save phyloseq object without primer-contaminated taxa

```{r}
saveRDS(d_16S, file =paste0(here::here("data/data_final"), "/d_16S.rds"))
saveRDS(d_pmoA, file =paste0(here::here("data/data_final"), "/d_pmoA.rds"))
saveRDS(d_mcrA, file =paste0(here::here("data/data_final"), "/d_mcrA.rds"))
```


## Sequences and clusters

### Final

```{r}
track_wkflow(list("16S"=d_16S,
                  "pmoA"=d_pmoA, 
                  "mcrA"=d_mcrA
                  ))
```

### Accross steps 


```{r}
rownames(track_16S) [rownames(track_16S) == "ASV denoising"] <- "ASV table"
rownames(track_16S) [rownames(track_16S) == "Paired sequences without chimera and longer than 200bp"] <- "Seq wo chimera (+200bp)"
rownames(track_mcrA) [rownames(track_mcrA) == "Forward sequences wo chimera (+200 bp)"] <- "Seq wo chimera (+200bp)"
rownames(track_pmoA) [rownames(track_pmoA) == "Forward sequences wo chimera (+200 bp)"] <- "Seq wo chimera (+200bp)"


library(tibble)
result <- bind_rows(
  list(
    "16S" = rownames_to_column(track_16S, "step"),
    "mcra" = rownames_to_column(track_mcrA, "step"),
    "pmoA" = rownames_to_column(track_pmoA, "step")
  ),
  .id = "id"
)

```

```{r}
result |>
  filter(step%in%intersect(rownames(track_16S), rownames(track_mcrA))) |>
  filter(step != "Paired sequences") |>
  filter(step != "Raw Forward sequences") |>
   mutate(step = factor(step, levels = c(rownames(track_16S)))) |>
  ggplot(aes(x = id, y = nb_sequences, fill = id)) +
  geom_bar(stat = "identity") +
  facet_grid(~step) + 
  no_legend()
```

```{r}
result |>
  filter(step%in%rownames(track_16S)[(rownames(track_16S) %in% rownames(track_mcrA))]) |>
  filter(step != "Raw Forward sequences") |>
    filter(step != "Paired sequences") |>
   mutate(step = factor(step, levels = c(rownames(track_16S)))) |>
  ggplot(aes(x = id, y = nb_clusters, fill = id)) +
  geom_bar(stat = "identity") +
  facet_grid(~step) + 
  no_legend()
```

### Accross step, by sample

```{r}
track_16S_samp  <- tar_read(track_by_samples, store = here::here("_targets/"))
track_mcrA_samp <- tar_read(track_by_samples, store = here::here("_targets_mcrA/"))
track_pmoA_samp <- tar_read(track_by_samples, store = here::here("_targets_pmoA/"))
```

```{r}
#| message: false
# Build a long data frame with one row per (marker, sample, step).
# track_by_samples is a named list (one element per sample) of data frames
# with rownames = step names and columns nb_sequences / nb_clusters / nb_samples.
samp_to_long <- function(track_samp, marker) {
  bind_rows(lapply(names(track_samp), function(s) {
    data.frame(
      marker = marker,
      sample = s,
      rownames_to_column(track_samp[[s]], "step"),
      check.names = FALSE
    )
  }))
}

result_samp <- bind_rows(
  samp_to_long(track_16S_samp,  "16S"),
  samp_to_long(track_mcrA_samp, "mcrA"),
  samp_to_long(track_pmoA_samp, "pmoA")
) |>
  mutate(nb_clusters = ifelse(nb_sequences==0, 0, nb_clusters)) |> 
  mutate(
    step = ifelse(step == "ASV denoising", "ASV table", step),
    step = ifelse(step == "ASV denoising", "ASV table", step),
    step = factor(step, levels = c(
      "ASV table",
      "OTU after vsearch reclustering at 97%",
      "OTU vs after mumu cleaning algorithm",
      "OTU vs + mumu + rarefaction by sequencing depth",
      "OTU vs + rarefaction by sequencing depth"
    ))
  ) |>
    filter(step %in%c( "ASV table",
      "OTU after vsearch reclustering at 97%",
      "OTU vs after mumu cleaning algorithm"))
```

```{r}
#| fig-width: 14
#| fig-height: 5
result_samp |>
  ggplot(aes(x = marker, y = nb_sequences, color = marker)) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_grid(~step) +
  no_legend() +
  theme_idest() +
  xlab("") +
  ylab("Sequences per sample")
```

```{r}
#| fig-width: 14
#| fig-height: 5
result_samp |>
  ggplot(aes(x = step, y = nb_clusters, color = step, group = sample)) +
  geom_line(alpha = 0.7, color="grey80", position = position_jitter(seed=2213,width = 0.2)) +
  geom_point(width = 0.2, size = 1, alpha = 0.5, position=position_jitter(seed = 2213, width=0.2)) +
  facet_grid(~marker) +
  theme_idest() +
   scale_x_discrete(labels = c("ASV table" = "ASV",
                              "OTU after vsearch reclustering at 97%" = "OTU",
                              "OTU vs after mumu cleaning algorithm" = "mumu")) +
  xlab("") +
  ylab("Clusters (OTUs) per sample") +
  no_legend() 
```

## OTU richness per sample

```{r}
richness_archaea <- data.frame(
  Marker = "Archaea 16S",
  Sample = sample_names(d_16S),
  OTUs = estimate_richness(d_16S, measures = "Observed")$Observed
)
richness_mcrA <- data.frame(
  Marker = "mcrA",
  Sample = sample_names(d_mcrA),
  OTUs = estimate_richness(d_mcrA, measures = "Observed")$Observed
)
richness_pmoA <- data.frame(
  Marker = "pmoA",
  Sample = sample_names(d_pmoA),
  OTUs = estimate_richness(d_pmoA, measures = "Observed")$Observed
)

richness_all <- rbind(richness_archaea, richness_mcrA, richness_pmoA)

ggplot(richness_all, aes(x = Marker, y = OTUs, fill = Marker)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 1.5, alpha = 0.6) +
  theme_idest() +
  xlab("") +
  ylab("OTU richness per sample") +
  theme(legend.position = "none")
```

## Read counts per sample

```{r}
reads_archaea <- data.frame(
  Marker = "Archaea 16S",
  Sample = sample_names(d_16S),
  Reads = sample_sums(d_16S)
)
reads_mcrA <- data.frame(
  Marker = "mcrA",
  Sample = sample_names(d_mcrA),
  Reads = sample_sums(d_mcrA)
)
reads_pmoA <- data.frame(
  Marker = "pmoA",
  Sample = sample_names(d_pmoA),
  Reads = sample_sums(d_pmoA)
)

reads_all <- rbind(reads_archaea, reads_mcrA, reads_pmoA)

ggplot(reads_all, aes(x = Marker, y = Reads, fill = Marker)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 1.5, alpha = 0.6) +
  theme_idest() +
  xlab("") +
  ylab("Read count per sample") +
  theme(legend.position = "none")
```

## Carbon footprint summary

```{r}
carbon_table <- data.frame(
  Marker = c("Archaea 16S", "mcrA", "pmoA"),
  gCO2_equiv = c(
    res_ga$carbon_footprint_total_gCO2,
    res_ga_mcrA$carbon_footprint_total_gCO2,
    res_ga_pmoA$carbon_footprint_total_gCO2
  )
)
kable(carbon_table, col.names = c("Marker", "Carbon footprint (gCO2 equiv.)"))
```

{{< include _session_info.qmd >}}
