---
title: "Bioinformatics pipeline summary"
subtitle: "Where we see the pipeline processes"
date: last-modified
author:
  - name: Adrien Taudière
    url: https://adrientaudiere.github.io/bioinfo.starter/analysis/01_bioinformatics.qmd
    affiliation: IdEst
    affiliation-url: https://orcid.org/my-orcid?orcid=0000-0003-1088-1182
citation: true
# bibliography: bibliography.bib
# link-citations: true
---

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(targets)
library(MiscMetabar)
here::i_am("analysis/01_bioinformatics.qmd")
source(here::here("R/functions.R"))
source(here::here("R/contamination.R"))
```

```{r child="../R/styles.qmd"}
```

# Summary of the bioinformatic pipeline

## Carbon footprint

```{r, message=FALSE, warning=FALSE}
res_ga <- greenAlgoR::ga_targets(store = here::here("_targets/"))
print(paste0(res_ga$carbon_footprint_total_gCO2, " gCO2 equiv. emitted for the whole pipeline"))
```

## Timeline and cpu usage

```{r, message=FALSE, warning=FALSE}
log_data <- autometric::log_read(
  here::here("data/data_final/autometric_log.txt"),
  units_memory = "gigabytes"
)

log_data |>
  filter(!grepl("conclude:", phase)) |>
  filter(!grepl("prepare:", phase)) |>
  filter(!grepl("__DEFAULT__", phase)) |>
  group_by(pid, phase) |>
  summarise(
    time = last(time) - first(time),
    cpu = mean(core),
    resident = mean(resident),
    .groups = "drop"
  ) |>
  group_by(phase) |>
  slice_max(time, n = 1, with_ties = FALSE) |>
  ungroup() |>
  ggplot(aes(x = time, y = reorder(phase, time), color = resident)) +
  geom_segment(aes(linewidth = cpu, xend = time, x = 0)) +
  scale_color_viridis_b("Memory (Gb)", end = 0.9, direction = -1) +
  geom_text(aes(label = paste0(round(time), "s/ ",round(resident), "Gb/",round(cpu), " cpu")), hjust = -0.1, size = 3, color="gray20") +
  theme_idest() +
  scale_x_log10() +
  xlab("Time (in seconds)") +
  ylab("")

```

## Main phyloseq object

You may want to select another targets such as d_asv or d_vs_mumu for example. 

```{r}
d_pq <- clean_pq(tar_read("d_vs", store=here::here("_targets/")))
```

```{r}
summary_plot_pq(d_pq)
```

```{r Visualise target plan, out.width='80%'}
tar_glimpse(script=here::here("_targets.R"), targets_only = TRUE, callr_arguments = list(show = FALSE))
```

```{r}
tar_meta(store=here::here("_targets/"), targets_only = TRUE) |> 
  dplyr::mutate(time = paste0(seconds %/% 3600,":",seconds %/% 60,":",floor(seconds %% 60)))|>
  dplyr::select(name, seconds, bytes, format, time) |>
  dplyr::mutate(Gb=round(bytes/10^9,2)) |>
  dplyr::arrange(desc(seconds), desc(bytes))  |> 
  kable()
```


## Load phyloseq object from targets store

```{r}
d_pq <- tar_read("d_vs", store=here::here("_targets/"))
```

<hr/>

The {targets} package is at the core of this project. Please read the intro of the [user manual](https://books.ropensci.org/targets/) if you don't know {targets}.

::: {.cell .page-columns .page-full .column-screen .grey-section}
The {targets} package store ... *targets* in a folder and can load (`tar_load()`) and read (`tar_read`) object from this folder.
:::

## Sample data

```{r}
DT::datatable(d_pq@sam_data)
```

## Sequences, samples and clusters across the pipeline


::: {.panel-tabset .nav-pills}
## All samples together

```{r}
track_sequences_samples_clusters <- tar_read(track_sequences_samples_clusters, store=here::here("_targets/"))
kable(track_sequences_samples_clusters)
```


```{r}
track_sequences_samples_clusters$step <- factor(rownames(track_sequences_samples_clusters), levels = rownames(track_sequences_samples_clusters))

scale_factor <- max(track_sequences_samples_clusters$nb_sequences) / max(track_sequences_samples_clusters$nb_clusters, na.rm = TRUE)

ggplot(track_sequences_samples_clusters, aes(x = step)) +
  geom_bar(aes(y = nb_sequences), stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_line(aes(y = nb_clusters * scale_factor, group = 1), color = "firebrick", linewidth = 1) +
  geom_point(aes(y = nb_clusters * scale_factor), color = "firebrick", size = 2) +
  scale_y_continuous(
    name = "Number of sequences",
    sec.axis = sec_axis(~ . / scale_factor, name = "Number of clusters")
  ) +
  theme_idest() +
  geom_text(aes(y = nb_sequences, label = paste0(nb_samples, " samp.")), vjust = -0.5, color = "grey20", size = 3) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y.left = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick")
  ) +
  xlab("")
```


## Per samples

```{r}
#| message: false
#| results: asis
tab_samp <- tar_read(track_by_samples, store=here::here("_targets/"))

for (li in names(tab_samp)) {
  print(knitr::kable(tab_samp[[li]], caption = li, format="html"))
  cat('\n<!-- -->\n\n')
}
```

## Violin plots

```{r}
#| message: false
#| warning: false
tab_all <- bind_rows(tab_samp, .id = "samples") |>
    tibble() 

tab_all$step <- factor(rep(rownames(tab_samp[[1]]), nrow(tab_all)/nrow(tab_samp[[1]])), levels=rownames(tab_samp[[1]]))

tab_all <- tab_all |>
  left_join(d_pq@sam_data, by = c("samples" = "samples_names_common"))

ggplot(tab_all, aes(x = step, y = log10(nb_sequences))) +
  geom_violin(fill = "grey80") +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1) +
  geom_line(aes(group = samples), color = "grey50", alpha = 0.1) +
  theme_idest() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Number of sequences (Log10)") +
  ylab("")


tab_all |>
  filter(nb_clusters > 0 & nb_clusters <1000) |>
ggplot(aes(x = step, y = nb_clusters)) +
  geom_violin(fill = "grey80") +
  geom_jitter(aes(color=sample_type),width = 0.15, alpha = 0.4, size = 1) +
  geom_line(aes(group = samples, color=sample_type), alpha = 0.1) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "pointrange",
    color = "grey20"
  ) +
  theme_idest() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Number of clusters") +
  ylab("")
```

:::


## Make krona html files

```{r}
krona(
  clean_pq(d_pq, simplify_taxo = TRUE),
  file = here::here("data/data_final/krona_nb_seq.html")
)
krona(
  clean_pq(d_pq, simplify_taxo = TRUE),
  nb_seq = FALSE,
  file =  here::here("data/data_final/krona_nb_taxa.html")
)
```


## Verify controls

```{r}
d_pq |>
   subset_samples(sample_type%in% c("blanc")) |>
   clean_pq() |>
   otu_table() |> data.frame() |> (\(x) log1p(x))() |>  pheatmap::pheatmap(cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, show_colnames = FALSE, color = colorRampPalette(c("white", "firebrick3"))(100)) 

d_pq@sam_data$neg_control <-
   d_pq@sam_data$sample_type %in% c("blanc") 
   
hill_pq(clean_pq(d_pq), fact = "neg_control", one_plot = TRUE) + theme_idest()
```

```{r}
devtools::load_all("~/Nextcloud/IdEst/Projets/arround_MiscMetabar/comparpq/")
devtools::load_all("~/Nextcloud/IdEst/Projets/arround_MiscMetabar/tidypq/")


estim_diff_pq(clean_pq(d_pq), fact = "neg_control")
```


#TODO c'est quoi sample_type = "ink"
Quels sont les taxons présents dans les blancs ?

```{r}
ggvenn_pq(d_pq, fact="neg_control")
```

### Classification des contaminations

```{r}
neg_ctrl_results <- classify_neg_control_taxa(
  d_pq,
  neg_control = d_pq@sam_data$sample_type %in% c("blanc")
)
```

```{r}
knitr::kable(neg_ctrl_results$classification |>
  dplyr::select(taxon, category, total_reads, reads_in_neg, reads_in_samples,
                n_neg_samples, n_non_neg_samples, ratio_non_neg_to_neg))
```

```{r}
neg_ctrl_results$summary_plot
```

```{r}
neg_ctrl_results$distribution_plot
```

```{r}
neg_ctrl_results$heatmap_by_category
```

```{r}
source("R/contamination.R")
classif_neg_control_taxa <- classify_neg_control_taxa(clean_pq(d_pq), neg_control = clean_pq(d_pq)@sam_data$sample_type %in% c("blanc"))

classif_neg_control_taxa$classification
classif_neg_control_taxa$summary_plot
classif_neg_control_taxa$distribution_plot 
classif_neg_control_taxa$heatmap_by_category
```





{{< include _session_info.qmd >}}
